---
title: "Assignment 11"
author: "Ethan Jantz"
date: "4/4/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Question 1

Write a loop that iterates over the numbers 1 - 10 and adds 5 to each before storing the results in a new vector called `output`. 

```{r}
output <- vector("numeric", 10)

for(i in 1:10) {
  output[i] <- i + 5
}

print(output)

output2 <- c(1:10 + 5) # Does the same thing

print(output2)
```

# Question 2

Create the matrix `mat_x` and perform the following operations:

a. Make a for-loop that calculates the sum for each column of the matrix. 
b. Use the apply function to perform the same operation.

```{r}
mat_x <- matrix(1:120, nrow = 20, ncol = 6)

for(i in seq_len(ncol(mat_x))) {
  print(
    paste("The sum of column", i, "in mat_x is", sum(mat_x[,i]), 
          sep = " ") # Ensures R includes a space between each entry
  )
}
```

```{r}
print(apply(mat_x, 2, 
            function(x) sum(x))) # This can perform the sum, 
# but I can't get it to print in a neat way as I did abote
```

# Question 3

Create a data frame with two numeric and one character vector. Write a loop that loops over the columns and reports the mean of the column if it is numeric and the number of observations if it's a character vector. 

```{r}
df <- data.frame( 
  V1 = 1:10,
  V2 = c(letters[1:10]),
  V3 = rchisq(10, df = 10)
)

for(i in seq_len(ncol(df))) {
    if (class(df[,i]) %in% c("integer", "numeric")) print(mean(df[,i]))
    if (class(df[,i]) == "character") print(length(df[,i]))
}
```

# Question 4

Using a loop, generate 10 random normals from distributions with means -10, 0, 10, and 100. Store them in a new matrix called `mat_distributions`. 

```{r}
means <- c(-10, 0, 10, 100)
mat_distributions <- matrix(nrow = 10, ncol = length(means))

for(i in seq_along(means)) {
  mat_distributions[,i] <- rnorm(nrow(mat_distributions), mean = means[i])
}

mat_distributions
```

# Question 5

use a simple `ifelse` statement to add a new column 'male.teen' to a dataframe. This is a dummy variables, indicating 1 ift he observation is a male younger than 20 and 0 otherwise. 

```{r}
respondent.df <- data.frame(
  name = c("Sue", "Eva", "Henry", "Jan", "Mary", "John"),
  sex = c("f", "f", "m", "m", "f", "m"),
  years = c(21,31,29,19, 23,33)
  )

respondent.df$male.teen <- ifelse(respondent.df$years < 20 & respondent.df$sex == "m", 
                                  1, 0)

print(respondent.df)
```

# Question 6

Using the same dataset, create a new column where values greater than 30 in the column `year` are coded as NA. Other values should stay the same. Use both `ifelse` and `if_else`.

```{r}
respondent.df$age <- ifelse(respondent.df$years > 30, NA, respondent.df$years)

respondent.df$age2 <- if_else(respondent.df$years > 30, NA, TRUE) # Not sure I understood the question here
```

# Question 7

Use `tapply` to calculate the average, minimum, and maximum number of years for female and male respondents. 

```{r}
tapply(respondent.df$years, respondent.df$sex, mean)

tapply(respondent.df$years, respondent.df$sex, min)

tapply(respondent.df$years, respondent.df$sex, max)
```

# Question 8

Create a loop that will number the publications for each author. "Tidy" the data so that each row represents on author only.

```{r}
data <- as_tibble(
  author = c("Author1", "Author1","Author2", "Author3", "Author3","Author3", "Author4", "Author5"), 
  pub = c("Pub1", "Pub2", "Pub3", "Pub4","Pub5", "Pub6", "Pub7", "Pub8"), 
  type = c("preprint", "article", "preprint","article", "article", "preprint", "preprint", "article")
)

for(i in seq_along(data$pub)) {
  
}
```