---
title: "Assignmen 11"
author: "Ethan Jantz"
date: "4/25/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warnings = FALSE)
library(kableExtra)
library(stevedata)
library(tidyverse)

data <- gss_wages
```

Create a nice summary table of the following:

* Average income for females
* Average income for males
* Number of children for females
* Number of children for males
* Difference in Income
* Difference in number of children

Must consider at least 7 or 8 years, max 15. You can select certain years (5yr avg) or a given time period (1990 - 2010). 

# Create Summary Table

I'll be summarizing the most recent 10 years in the data, 2000 - 2018, for this project. The first step is to create the summary table.

```{r}
table <- data %>%
  group_by(gender, year) %>% # We want summaries for both year and gender
  summarize(avg_income = mean(realrinc, na.rm = T),
            children = mean(childs, na.rm = T)) %>%
  ungroup() %>% # Because i don't want groups to mess anything up
  pivot_wider(names_from = gender, # We want separate columns for Male and Female in the summarized vars
              values_from = c(avg_income, children)) %>%
  mutate(year = as.character(year),
         avg_income_diff = avg_income_Male - avg_income_Female,
         children_diff = children_Male - children_Female,
         income_diff_change = avg_income_diff - lag(avg_income_diff)) %>% # Note, the directionality of diff_ is Male to Female
  filter(year %in% 2000:2018) %>% # This creates the final table 
  select(year,
         avg_income_Male, children_Male,
         avg_income_Female, children_Female,
         avg_income_diff, children_diff,
         income_diff_change) # reorder variables for table
```

# Create Kable Table

Then I want to style the table.

```{r}
table %>%
  # basic table arrangements, label aesthetics
  kable(align = "lrrrrrrr", 
        col.names = c("Year[note]", rep(c("Income", "Children"), 3), "Income Change[note]"),
        caption = "Average Income and # Children by Year, 2008 - 2018", 
        digits = c(0, rep(2, 6)),
        format.args = list(big.mark = ",")) %>%
  # full-table aesthetics
  kable_styling(font_size = 20,
                html_font = "Lucida",
                full_width = F) %>%
  # header stuff
  row_spec(0, bold = T, background = "#e5e5e5") %>%
  add_header_above(c(" " = 1, "Male" = 2, "Female" = 2, "Difference[note]" = 3)) %>%
  # highlights
  row_spec(which(table$avg_income_diff %in% max(table$avg_income_diff)), color = "red") %>% # first, color the rows I want to highlight
  column_spec(c(1,2,3,4,5,6,8), # everything except the difference cols
              color = "black")
  
  # footnotes
  add_footnote(c("Only years between 2000 and 2018 were included", 
                 "Change from previous year", 
                 "Differences were calculated subtracting female averages from male averges"))
```

