---
title: "Assignment 6"
author: "Ethan Jantz"
date: "2/20/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)

data <- read_csv(here::here("HW 6", "api_data.csv"))
```

# Step 1

## Question 1: Unit of Analysis

The unit of analysis is the school-characteristic. Each row represents one school and an observation of its student population. There are three population characteristics observed in this data, `meals`, `colgrad`, and `fullqual`. 

## Question 2: Tidiness

The data is messy, but not incredibly so. The `variable_name` and `percentage` columns are tidy, but the `community_schooltype` column is not. `community_schooltype` will beak out into two variables, `community` and `schooltype`, using `separate()`. The final shape of the data will have 30,000 rows and 7 columns. I will also convert character-type variables to factors for ease of analysis later on.


```{r tidy}
data <- data %>%
  separate(col = community_schooltype,
           into = c("community", "schooltype"),
           sep = "_") %>%
  mutate(across(.cols = where(is.character), .fns = as.factor))

glimpse(data)
```

## Question 3: Missing data

Let's use `dplyr::summarize` to see if there is any missing data. 

```{r missingness}
data %>%
  summarize(across(.cols = everything(),
                   .fns = ~ sum(is.na(.x))))
```

It looks like there are no missing observations in this dataset. This makes sense, considering it's simulated data.

# Step 2

## Question 1: Visualizing Descriptive Statistics

Let's take a look at the frequency of the variables in the data using density plots for continuous variables and bar plots for categorical variables. I will not be exploring `uid` or `county` as per the instructions. 

```{r density plots}
plot_dens <- data %>%
  ggplot(aes(percentage, fill = variable_name, color = variable_name)) +
  geom_density(alpha = .1)
```

```{r bar plots}

```