---
title: "Assignment 12"
author: "Ethan Jantz"
date: "4/10/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)

state_names <- state.name
sentences <- as_tibble(sentences) %>%
  rename(sentence = value)
```

# Part 1

Practicing string extraction with the `state.name` dataset.

## Q1

Extract Arizona, California, Illinois, and Oregon from the data.

```{r}
pattern <- "Arizona|California|Illinois|Oregon"
selected <- str_extract(state_names, pattern) %>% 
  na.omit() # Remove the NA values for non-matches

paste(selected)
```

## Q2

Extract all states that begin with the letter 'a'. Use both `str_subset` and `str_extract` to see how the syntax differs. In both cases, extract the full name of all the states.

### `str_extract`

```{r}
pattern <- "^[A].*"
selected <- str_extract(state_names, pattern) %>%
  na.omit()

paste(selected)
```

### `str_subset`

```{r}
# same pattern as above
selected <- str_subset(state_names, pattern) # Doesn't require removal of NA values
paste(selected)
```

## Q3

All states with names ending in a or e. 

```{r}
pattern <- ".*(a|e)$"
selected <- str_subset(state_names, pattern)
paste(selected)
```

## Q4

All states with two words (such as North Dakota).

```{r}
pattern <- "\\s"
selected <- str_subset(state_names, pattern)
paste(selected)
```

## Q5

All states with n, t, w, or c at any point.

```{r}
pattern <- "[ntwc]"
selected <- str_subset(state_names, pattern)
paste(selected)
```

## Q6

Count the states without w, z, or y.

```{r}
pattern <- "[wzyWZY]"
selected <- str_subset(state_names, pattern, negate = T)
length(selected)
```

## Q7

Extract states with at least one c or i, excluding where these letters are the first letter.

```{r}
pattern <- "^[^CI][ci]+"
selected <- str_subset(state_names, pattern)
paste(selected)
```

## Q8

State names that are exactly 6 letters long 

```{r}
pattern <- "^.{6}$"
selected <- str_subset(state_names, pattern)
paste(selected)
```

AND at least 6 letters long.

```{r}
pattern <- "^.{6,}$"
selected <- str_subset(state_names, pattern)
paste(selected)
```

## Q9

Match any string in the `state_name` vector.

```{r}
pattern <- ".*"
selected <- str_subset(state_names, pattern)
paste(selected)
```

## Q10

Extract all states that start with two consonants.

```{r}
pattern <- "^[^AEIOUaeiou]{2}"
selected <- str_subset(state_names, pattern)
paste(selected)
```

## Q11

Extract all states with two vowels in a row.

```{r}
pattern <- "[AEIOUaeiou]{2,}"
selected <- str_subset(state_names, pattern)
paste(selected)
```

## Q12

Extract states that start with a vowel and end with a consonant.

```{r}
pattern <- "^[AEIOU].*[^aeiou]$"
selected <- str_subset(state_names, pattern)
paste(selected)
```

# Part 2

Create new columns in the `sentences` tibble using dplyr. The columns should be as follows:

1. The number of letters in each sentence.

2. A dummy variable equal to 1 if the sentence contains a word ending with "ing", and 0 otherwise. 

3. A dummy variable equal to 1 if the sentence contains a color such as red, orange, yellow, green, blue, pink, and purple. 

4. The number of words in each sentence. 

```{r}
colors <- c("orange", "red", "yellow", "green", "blue", "pink", "purple", "black", "white")
sentences <- sentences %>%
  mutate(
    letter_count = str_count(sentence, "[A-Za-z]"),
    has_ing = if_else(str_detect(sentence, "ing\\b"), # \\b denotes the edge of a word
                      1, 0),
    has_color = if_else(str_detect(sentence, 
                                   paste0("(?i)\\b(", paste0(colors, collapse = "|"), ")\\b")),
                        1, 0),
    word_count = str_count(sentence, boundary("word")) # boundary() is a modifier for stringr functions
    # in this case the boundary is words, but you could make it lines, characters, or sentences
  )
```

Find the following:

1. The average number of letters in across sentences.

2. The number of words ending in ing in the data.

3. The number of times colors are mentioned in the data.

```{r}
sentences %>%
  summarize(
    letters_avg = mean(letter_count),
    ing_sum = sum(has_ing),
    color_sum = sum(has_color)
  )
```


